A game theoretic model of the behavioural gaming that takes place at the EMS - ED interface\tnoteref{t1}

### 2022-04-05 sections/conclusion/main.tex
Apply changes to conclusion

The motivation behind this study has been that emergency departments are
under a lot of pressure to treat patients.
This is in practice often centrally controlled through a mechanism of some sort
of performance measure target.
This paper shows how this can negatively impact the pathway of both the
ambulance patients and the ambulance service itself.
Due to some managerial decision making that takes place at the ED, ambulances
stay blocked outside of the ED at the hospital's parking zone in an attempt
to satisfy these regulations.
The main contributions of this paper are:
\begin{itemize}
    \item A queueing model with 2 consecutive waiting spaces where one would
    serve as a parking space for the ambulances;
    \item Analytic performance measure formulas for the queueing model;
    \item A 3-player game theoretic model between the EMS and two EDs;
    \item Numerical experiments showing emergent behaviour of gaming between
    EDs and the EMS.
\end{itemize}
Although our research is motivated by the particular EMS-ED example, our 
developed modelling framework and behavioural insights has application to 
similar systems across a range of sectors and settings.
The queueing model can be applied to any setting where individuals may be
blocked on a separate queue.
An example of such setting can be any type of delivery service where customers
can purchase goods either online or walk in the store.
At busier times, the person delivering the food may be blocked outside the
restaurant in an attempt to improve the waiting times for walk-in customers.

This study explores a generic 3-player game theoretic model between the
decision makers of two queueing systems and a service that distributes
individuals to these two systems (section \ref{sec:model_overview}).
It also describes the construction of the underlying queueing theoretic model
that has a tandem buffer and a single service centre (section
\ref{sec:queueing_model}).
Furthermore, the formulas for the performance measures of the queueing model
are also derived (sections \ref{sec:waiting_time}, \ref{sec:blocking_time},
\ref{sec:proportion_within_target}).
This novel queueing model is the first contribution of the paper.
The game theoretic model is then applied to a healthcare scenario by looking at
the interface between the EDs and the EMS.
The inefficiencies that emerge from the perspective of the EMS were explored
along with ways to apply some incentive mechanisms to improve them.
The key findings from this paper that were observed when playing the game
between two EDs and the EMS are:
\begin{itemize}
    \item Inefficiencies can be learned and emerge naturally
    \item Targeted incentivisation of behaviours can help escape inefficiencies
\end{itemize}
The former relates to the results of asymmetric replicator dynamics that showed
that inefficient scenarios can arise by playing the game while the latter
implies that the learned inefficiencies can be escaped by carefully applying
certain incentives to the players.
This applied game theoretic model is the second main contribution of this paper.

The model presented here assumes the presence of only two players that can
receive individuals.
However, in a realistic healthcare scenario an ambulance may have to decide
among multiple EDs.
An immediate extension of this work would be to consider a multiplayer system
that could represent a group of hospitals in a concentrated area.
Additionally, the game theoretic model that was created uses a discrete
strategy space for the EDs (something that is also present in various related
literature~\cite{deo2011centralized, knight2017measuring}).
The single threshold parameter that is used for the ED's decision may not be
the best way to describe the model.
In reality ED managers could have far more complex parameters for their
decision making process.
Moreover this work assumes that the EMS and EDs act in a selfish and rational
way by only aiming to satisfy their own objectives.
In fact, modelling behaviour in a healthcare setting is a much more complex
task where some cooperation may often be observed.
Another extension would be to explore the behaviour of the ED staff via an
agent-based model.
This in turn can be used to model emergent behaviour based on assumptions of
individual behavioural traits of ED staff.
Finally, future work can touch upon the derivation of a closed form formula for
the steady state probability vector of the queueing model
(section \ref{sec:queueing_model}) to allow for faster computations of \(\pi\).

>>>>

The motivation behind this study has been that emergency departments are
under a lot of pressure to treat patients.
This is, in practice, often centrally controlled through a mechanism of some
sort of performance measure target.
This paper shows how this can negatively impact the pathway of both the
ambulance patients and the ambulance service itself.
Due to some managerial decision making that takes place at the ED, ambulances
stay blocked outside of the ED at the hospital's parking zone in an attempt
to satisfy these regulations.
The main contributions of this paper are:
\begin{itemize}
    \item A queueing model with 2 consecutive waiting spaces where one would
    serve as a parking space for the ambulances;
    \item Analytic performance measure formulas for the queueing model;
    \item A 3-player game theoretic model between the EMS and two EDs;
    \item Numerical experiments showing emergent behaviour of gaming between
    EDs and the EMS.
\end{itemize}
Although our research is motivated by the particular EMS-ED example, our 
developed modelling framework and behavioural insights has application to 
similar systems across a range of sectors and settings.
The queueing model can be applied to any setting where individuals may be
blocked on a separate queue.
An example of such setting can be any type of delivery service where customers
can purchase goods either online or in-person.
At busier times, the person delivering the product may be blocked outside the
store in an attempt to improve the waiting times for walk-in customers.

This study explores a generic 3-player game theoretic model between the
decision makers of two queueing systems and a service that distributes
individuals to these two systems (section \ref{sec:model_overview}).
It also describes the construction of the underlying queueing theoretic model
that has a tandem buffer and a single service centre (section
\ref{sec:queueing_model}).
Furthermore, the formulas for the performance measures of the queueing model
are also derived (sections \ref{sec:waiting_time}, \ref{sec:blocking_time},
\ref{sec:proportion_within_target}).
This novel queueing model is the first contribution of the paper.
The game theoretic model is then applied to a healthcare scenario by looking at
the interface between the EDs and the EMS.
The inefficiencies that emerge from the perspective of the EMS were explored
along with ways to apply some incentive mechanisms to improve them.
The key findings from this paper that were observed when playing the game
between two EDs and the EMS are:
\begin{itemize}
    \item Inefficiencies can be learned and emerge naturally;
    \item Targeted incentivisation of behaviours can help escape inefficiencies.
\end{itemize}
The former relates to the results of asymmetric replicator dynamics that showed
that inefficient scenarios can arise by playing the game, while the latter
implies that the learned inefficiencies can be escaped by carefully applying
certain incentives to the players.
This applied game theoretic model is the second main contribution of this paper.

The model presented here assumes the presence of only two players that can
receive individuals.
However, in a realistic healthcare scenario an ambulance may have to decide
among multiple EDs.
An immediate extension of this work would be to consider a multiplayer system
that could represent a group of hospitals in a concentrated area.
Additionally, the game theoretic model that was created uses a discrete
strategy space for the EDs (something that is also present in various related
literature~\cite{deo2011centralized, knight2017measuring}).
The single threshold parameter that is used for the ED's decision may not be
the best way to describe the model.
In reality ED managers might adopt far more complex parameters for their
decision making process.
Moreover this work assumes that the EMS and EDs act in a selfish and rational
way by only aiming to satisfy their own objectives.
In some settings, cooperation may be observed and would therefore require an
adapted modelling approach.
Another extension would be to explore the behaviour of the ED staff via an
agent-based simulation model.
This in turn can be used to model emergent behaviour based on assumptions of
individual behavioural traits of ED staff.
Finally, future work could touch upon the derivation of a closed form formula for
the steady state probability vector of the queueing model
(section \ref{sec:queueing_model}) to allow for faster computations of \(\pi\).
